% ── Acknowledgement ────────────────────────────────────────────
\chapter*{Acknowledgement}
\addcontentsline{toc}{chapter}{Acknowledgement}

Completing this research project would not have been possible without the guidance and encouragement of several people, and I would like to take a moment to express my gratitude.

First and foremost, I wish to sincerely thank my supervisor, Mr.\ Ravi Supunya, whose continuous guidance and constructive feedback steered this project in the right direction from the very beginning. His willingness to discuss ideas, review progress, and challenge my assumptions helped me grow both as a developer and as a researcher.

I am also grateful to Sri Lanka Institute of Information Technology for providing the academic environment and resources that made this work possible. The lecturers and staff of the Department of Information Technology shared their expertise generously throughout my undergraduate years, and that knowledge has been instrumental in shaping this research.

I owe a great deal to my project group members, who worked alongside me on different components of the RubberIntelligence platform. The collaborative spirit of our team made it easier to tackle the integration challenges that inevitably arise in a multi-module system.

A special word of thanks goes to the rubber plantation owners and agricultural officers who took the time to share their practical knowledge about diseases, pests, and cultivation techniques. Their domain expertise directly influenced the design of the chatbot knowledge base and the selection of detectable threat classes.

Finally, I want to acknowledge my family and friends for their unwavering support and patience throughout this journey. Their encouragement, especially during the more demanding phases of development and testing, kept me motivated to see this through to the end.

\newpage
