% ════════════════════════════════════════════════════════════════
%  REFERENCE LIST (IEEE Style)
% ════════════════════════════════════════════════════════════════
\begin{thebibliography}{99}
\addcontentsline{toc}{chapter}{Reference List}

\bibitem{ref1}
Rubber Development Department of Sri Lanka, ``Annual Review of the Rubber Industry,'' Ministry of Plantation Industries, Colombo, 2023.

\bibitem{ref2}
M. Jayasuriya and K. Wijesuriya, ``Impact of leaf diseases on rubber yield: A decade of field observations in the wet zone of Sri Lanka,'' \textit{Journal of the Rubber Research Institute of Sri Lanka}, vol.~98, pp.~45--62, 2021.

\bibitem{ref3}
K. Perera, ``Challenges in agricultural extension services for smallholder rubber farmers in Sri Lanka,'' \textit{Tropical Agricultural Research}, vol.~31, no.~2, pp.~112--125, 2020.

\bibitem{ref4}
Rubber Development Department of Sri Lanka, ``Statistical Pocket Book of the Rubber Industry,'' 2022.

\bibitem{ref5}
R. De Silva and L. Fernando, ``Climate change and emerging disease patterns in Sri Lankan rubber plantations,'' \textit{Journal of Plant Pathology}, vol.~104, pp.~789--801, 2022.

\bibitem{ref6}
Telecommunications Regulatory Commission of Sri Lanka, ``Statistical Overview of Telecommunication Sector: Annual Report,'' 2023.

\bibitem{ref7}
S. P. Mohanty, D. P. Hughes, and M. Salath\'{e}, ``Using deep learning for image-based plant disease detection,'' \textit{Frontiers in Plant Science}, vol.~7, p.~1419, 2016.

\bibitem{ref8}
K. P. Ferentinos, ``Deep learning models for plant disease detection and diagnosis,'' \textit{Computers and Electronics in Agriculture}, vol.~145, pp.~311--318, 2018.

\bibitem{ref9}
S. Sladojevic, M. Arsenovic, A. Anderla, D. Culibrk, and D. Stefanovic, ``Deep neural networks based recognition of plant diseases by leaf image classification,'' \textit{Computational Intelligence and Neuroscience}, vol.~2016, p.~3289801, 2016.

\bibitem{ref10}
J. G. A. Barbedo, ``Plant disease identification from individual lesions and spots using deep learning,'' \textit{Biosystems Engineering}, vol.~180, pp.~96--107, 2019.

\bibitem{ref11}
I. P. Tetila \textit{et al.}, ``Identification of soybean foliar diseases using unmanned aerial vehicle images,'' \textit{IEEE Geoscience and Remote Sensing Letters}, vol.~14, no.~12, pp.~2190--2194, 2017.

\bibitem{ref12}
K. Thenmozhi and U. S. Reddy, ``Crop pest classification based on deep convolutional neural network and transfer learning,'' \textit{Computers and Electronics in Agriculture}, vol.~164, p.~104906, 2019.

\bibitem{ref13}
Y. Li, H. Wang, L. M. Dang, A. Sadeghi-Niaraki, and H. Moon, ``Crop pest recognition in natural scenes using convolutional neural networks,'' \textit{Computers and Electronics in Agriculture}, vol.~169, p.~105174, 2020.

\bibitem{ref14}
A. Wang, W. Zhang, and X. Wei, ``A review on weed detection using ground-based machine vision and image processing techniques,'' \textit{Computers and Electronics in Agriculture}, vol.~158, pp.~226--240, 2019.

\bibitem{ref15}
P. Bonnet \textit{et al.}, ``Plant identification: Experts vs.\ machines in the era of deep learning,'' \textit{Multimedia Tools and Applications}, vol.~79, pp.~31089--31117, 2020.

\bibitem{ref16}
K. Jha, A. Doshi, P. Patel, and M. Shah, ``A comprehensive review on automation in agriculture using artificial intelligence,'' \textit{Artificial Intelligence in Agriculture}, vol.~2, pp.~1--12, 2019.

\bibitem{ref17}
P. Lewis \textit{et al.}, ``Retrieval-augmented generation for knowledge-intensive NLP tasks,'' in \textit{Advances in Neural Information Processing Systems (NeurIPS)}, vol.~33, pp.~9459--9474, 2020.

\bibitem{ref18}
N. Reimers and I. Gurevych, ``Sentence-BERT: Sentence embeddings using Siamese BERT-networks,'' in \textit{Proc. EMNLP-IJCNLP}, 2019, pp.~3982--3992.

\bibitem{ref19}
J. Johnson, M. Douze, and H. J\'{e}gou, ``Billion-scale similarity search with GPUs,'' \textit{IEEE Transactions on Big Data}, vol.~7, no.~3, pp.~535--547, 2021.

\bibitem{ref20}
A. Ramcharan \textit{et al.}, ``Deep learning for image-based cassava disease detection,'' \textit{Frontiers in Plant Science}, vol.~8, p.~1852, 2017.

\end{thebibliography}

\newpage
